---
title: "ts_analysis"
author: "Bruno Henrique Schock"
date: "April 28, 2019"
output: html_document
---

## Univariate time series analysis

Hello fellas, here's a small "tutorial" on how to run a simple analysis on some univariate time series, since the analysis of it's kind, such as observing which data we are messing up, to the prediction analysis with some popular metrics used in the literature, I really hope it suits well for you.

At first, of course, we need to check, and if not, install our packages, I'm sure some of them you are very used to, but to keep sure I'll leave some information on what they are influencing in our analysis.


```{r setup, include=T}
#Do we have all the packages? If not, install it!
packages <- c("urca", "forecast", "curl", "tseries", "MAPA")
new.ones <- packages[!(packages %in% installed.packages()[,"Package"])]
if(length(new.ones)) install.packages(new.ones)

#Load them
library(urca)
library(forecast)
library(curl)
library(tseries)
library(MAPA)
```




```{r cars}
#Let's organize our data
orgdata <- list(nhtemp, austres, AirPassengers, EuStockMarkets[,4])
data <- list(as.matrix(orgdata[[1]]), as.matrix(orgdata[[2]]), as.matrix(orgdata[[3]]), as.matrix(orgdata[[4]]))
freq <- list(frequency(orgdata[[1]]), frequency(orgdata[[2]]), frequency(orgdata[[3]]), frequency(orgdata[[4]]))


#We need to peek our data too
par(mfrow=c(ifelse(length(data)%%2==F, length(data)/2, (length(data)-1)), ifelse(length(data)%%2==F, length(data)/2, (length(data)-1))))
for(i in 1:length(data))plot(ts(data[[i]], start = start(orgdata[[i]]), end=end(orgdata[[i]]), frequency = frequency(orgdata[[i]])))
```

## Including Plots

```{r pressure, echo=FALSE}
#Create our statistics matrix in some organized way
teststats <- matrix(NA, nrow=ncol(data[[1]]), ncol=6)
teststat <- list()
colnames(teststats) <- c('tau3', 'phi2', 'phi3', 'tau2', 'phi1', 'tau1')
rownames(teststats) <- colnames(data[[1]])

#Now our critical values matrices
cvals <- matrix(NA, ncol=3, nrow=6)
cval <- list()
colnames(cvals) <- c('1pct', '5pct', '10pct')
rownames(cvals) <- c('tau3', 'phi2', 'phi3', 'tau2', 'phi1', 'tau1')

for(j in 1:length(data)){
for(i in 1:ncol(data[[j]])){
  teststats[i,1:3] <- ur.df(data[[j]][,i], type='trend', selectlags = 'AIC')@teststat
  cvals[1:3,] <- ur.df(data[[j]][,1], type='trend', selectlags = 'AIC')@cval
  teststats[i,4:5] <- ur.df(data[[j]][,i], type='drift', selectlags = 'AIC')@teststat
  cvals[4:5,] <- ur.df(data[[j]][,1], type='drift', selectlags = 'AIC')@cval
  teststats[i,6] <- ur.df(data[[j]][,i], type='none', selectlags = 'AIC')@teststat
  cvals[6,] <- ur.df(data[[j]][,1], type='none', selectlags = 'AIC')@cval
}
teststat[[j]] <-teststats 
cval[[j]] <- cvals
}

```

```{r}
#More brick breeaking 
results <- matrix(NA, ncol=1, nrow=length(data))
colnames(results) <- 'Process kind:'
rownames(results) <- colnames(data)


#Now we test our conditions suck as sugested in Pfaff(2008)
for(j in 1:length(data)){
for(i in 1:ncol(data[[j]])){
  if(teststat[[j]][i,1] < cval[[j]][1,1] | teststat[[j]][i,1] < cval[[j]][1,2] |
     teststat[[j]][i,1] < cval[[j]][1,3]){
    results[j,] <- "Stationary around a linear tendency"
  }
  else{if(teststat[[j]][i,3] > cval[[j]][3,1] | teststat[[j]][i,3] > cval[[j]][3,2] |
          teststat[[j]][i,3] > cval[[j]][3,3]){
    results[j,] <- "Random walk with linear tendency"
  }
    else { if(teststat[[j]][i,4] < cval[[j]][4,1] | teststat[[j]][i,4] < cval[[j]][4,2] |
              teststat[[j]][i,4] < cval[[j]][4,3]){
      results[j,] <- "Stationary around a constant"
    }
      else { if(teststat[[j]][i,5] > cval[[j]][5,1] | teststat[[j]][i,5] > cval[[j]][5,2] |
                teststat[[j]][i,5] > cval[[j]][5,3]){
        results[j,] <- "Random walk with drift"
      }
        else { if(teststat[[j]][i,6] < cval[[j]][6,1] | teststat[[j]][i,6] < cval[[j]][6,2] |
                  teststat[[j]][i,6] < cval[[j]][6,3]){
          results[j,] <- "Stationary with zero mean"
        } else {
          results[j,] <- "Random walk without drift"
        }}}}}
  if (freq[[j]] == 1) {
    results[j,] <- results[j,]
  }
  else if ((nsdiffs(ts(data[[j]][,i], frequency=freq[[j]])) >= 1)) {
  results[j,] <- paste0(results[j,]," with seasonality")}
print(results[j,])
}
 print("Voil?!")
   }

```




```{r}
#Criamos as fun??es que far?o parte da an?lise
pred.perf = function(pred, test) {
  fitt = pred - test
  MAE = sum(abs(fitt))/length(test)
  MAPE = sum(abs(fitt/pred))/length(test)
  RSS = sum(fitt^2)
  MSE = RSS/length(test)
  RMSE = sqrt(MSE)
  perf = data.frame(MAE, MAPE, RSS, MSE, RMSE)
  }

#Criamos a fun??o
forefun <- function(train_pe, test_pe, plot.res=T, seriesname, lims=F, beg, fin, ybeg, yfin){
  
  #Aggregate forecasting methods
  autoarim <- auto.arima(train_pe)
  autoholt <- ifelse(frequency(train_pe)==1,
  HoltWinters(train_pe,gamma=F), HoltWinters(train_pe))
  autotbat <- tbats(train_pe)
  automapa <- mapaest(train_pe)

  #Predict
  predarim <- predict(autoarim, n.ahead=length(test_pe))
  predholt <- forecast(autoholt[[1]], h=length(test_pe))
  predtbat <- forecast(autotbat, h=length(test_pe))
  predmapa <- mapafor(train_pe, automapa, fh=length(test_pe))


  #Measure the performance
  arimperf <- pred.perf(as.numeric(predarim$pred), test_pe)
  holtperf <- pred.perf(as.numeric(predholt$forecast$xhat$mean), test_pe)
  tbatperf <- pred.perf(as.numeric(predtbat$mean), test_pe)
  mapaperf <- pred.perf(predmapa$outfor, test_pe)

  performance <- rbind(arimperf, holtperf, tbatperf, mapaperf)
  rownames(performance) <- c("SARIMA", "HoltWinters", "TBATs", "MAPA")
  

  if(plot.res == T & lims == F) {
    ts.plot(train_pe, predarim$pred, predholt$forecast$xhat$mean , predtbat$mean, ts(predmapa$outfor, start = start(test_pe), end=end(test_pe),
    frequency=frequency(train_pe)), col = 1:5, lty=1:5)
    legend("topleft", c(seriesname, "AutoArima", "HoltWinters", "TBATs", "MAPA"), lty = 1:5, 
           col = 1:5)
  }
  
  if(plot.res == T & lims == T) {
    ts.plot(train_pe, predarim$pred, predholt$forecast$xhat$mean , predtbat$mean, ts(predmapa$outfor, start = start(test_pe), end=end(test_pe),
            frequency=frequency(train_pe)), col = 1:5, lty=1:5, xlim=c(beg, fin), ylim=c(ybeg, yfin))
    legend("topleft", c(seriesname, "AutoArima", "HoltWinters", "TBATs", "MAPA"), lty = 1:5, 
           col = 1:5)
  }
  
    print(seriesname)
  print(performance[order(performance$MAPE),])

  }

```


```{r}

train_pe <- window(orgdata[[1]], start=start(orgdata[[1]]), end=c(1959,1))
test_pe <- window(orgdata[[1]], start=c(1959,1), end=end(orgdata[[1]]))
forefun(train_pe, test_pe, "", plot.res = T, lims=F)

```


```{r}
train_pe <- window(orgdata[[2]], start=start(orgdata[[2]]), end=c(1991,1))
test_pe <- window(orgdata[[2]], start=c(1991,1), end=end(orgdata[[2]]))
nhpres <- forefun(train_pe, test_pe, "", plot.res = T, lims=F)

```


```{r}
train_pe <- window(orgdata[[3]], start=start(orgdata[[3]]), end=c(1957,12))
test_pe <- window(orgdata[[3]], start=c(1958,1), end=end(orgdata[[3]]))

nhpres <- forefun(train_pe, test_pe, "", plot.res = T, lims=F)

```

